---
title: "Introduction-Basic Mapping"
author: "Kevin Stierhoff"
date: '`r format(Sys.time(), format = "%F")`'
output: ioslides_presentation
logo: ../Images/noaa-logo.png
---

```{r setup, include=FALSE}
# Install and load pacman (library management package)
if (!require("pacman")) install.packages("pacman")

# Install and load required packages from CRAN ----
pacman::p_load(knitr,tidyverse,here,emo,bookdown)

knitr::opts_chunk$set(message = F, fig.align = 'center')
```

## Outline

+ What makes a map?
+ Let's make a map!
+ Base graphics
+ ggplot2

## What is a map?

_**According to Wikipedia:**_

A map is a symbolic depiction emphasizing relationships between elements of some space, such as objects, regions, or themes.  

Many maps are static, fixed to paper or some other durable medium, while others are dynamic or interactive.  

## Let's make a map

Load some transect data from a survey of coastal pelagic fishes (CPS) off the west coast of the U.S. and Canada

```{r,echo=T}
# Read data file
cps <- read_csv(here("Data/cps.csv")) 

# Show first 10 rows
head(cps, 5)
```  

## Let's make a map

Create a **VERY** basic map using base graphics (`r emo::ji("sick")`)

```{r,echo=T}
# Plot lat/long
plot(cps$long, cps$lat)
```  

## Let's make a (slightly better) map

The same **VERY** basic map using ggplot2 (still `r emo::ji("sick")`)

```{r,echo=T}
# Plot lat/long
ggplot(cps, aes(long, lat)) + geom_point()
``` 

## Let's make a (barely acceptable) map

The same **VERY** basic ggplot2 map with some tweaks (still `r emo::ji("sick")`)

```{r,echo=T}
# Plot lat/long
ggplot(cps, aes(long, lat)) + geom_point() + coord_map() + theme_bw()
```
